Reconstructing Propositional Proofs in Type Theory

### 2017-10-27 paper/sections/introduction.tex
Fixed typo.
Cosmetic changes.

This paper has been organized in the following way.
In Section~\ref{sec:type-theory} some limitations of type theory
are discussed from proof-reconstruction point view.
In Section~\ref{sec:metis-language-and-proofs}, we introduce the
\Metis prover.
In Section~\ref{sec:proof-reconstruction} we show our
approach to reconstruct \Metis proofs.
Related work is described in Section~\ref{sec:related-work}.
Conclusions and suggestions for future work are  presented in
Section~\ref{sec:conclusions}.

@@

The source codes accompanying this paper (programs, libraries, and
examples) are available as repositories at \verb!GitHub!:

\begin{itemize}
  \item The \Athena program that translates proofs generated by \Metis to \Agda:
  \url{http://github.com/jonaprieto/athena}.
  \item The \verb!agda-prop! library as a formalization in \Agda for
    classical propositional logic:
  \url{http://github.com/jonaprieto/agda-prop}.
\item The \verb!agda-metis! library as a formalization in \Agda to
  justify \Metis proofs of classical propositional logic:
  \url{http://github.com/jonaprieto/agda-prop}.
\end{itemize}

The proof-reconstruction was carried on for \Metis~2.3
(release~20170822), and the accompanying source code was tested with
\Agda~2.5.3, \Agda standard library~0.14 and \prg{GHC}~8.2.1.

\end{document}

>>>>

This paper has been organized in the following way.  In
Section~\ref{sec:type-theory}, some limitations of type theory are
discussed from proof-reconstruction point view.  In
Section~\ref{sec:metis-language-and-proofs}, we introduce the \Metis
prover.  In Section~\ref{sec:proof-reconstruction}, we show our
approach to reconstruct \Metis proofs.  Related work is described in
Section~\ref{sec:related-work}.  Conclusions and suggestions for
future work are presented in Section~\ref{sec:conclusions}.

@@

The source codes accompanying this paper (programs, libraries, and
examples) are available at \verb!GitHub!:
\begin{itemize}
  \item The \Athena program that translates proofs generated by \Metis to \Agda:
  \url{http://github.com/jonaprieto/athena}.
  \item The \verb!agda-prop! library as a formalization in \Agda for
    classical propositional logic:
  \url{http://github.com/jonaprieto/agda-prop}.
\item The \verb!agda-metis! library as a formalization in \Agda to
  justify \Metis proofs of classical propositional logic:
  \url{http://github.com/jonaprieto/agda-prop}.
\end{itemize}
The proof-reconstruction was carried on for \Metis~2.3
(release~20170822), and the programs and libraries were tested with
\Agda~2.5.3, \Agda standard library~0.14, and \prg{GHC}~8.2.1.
\end{document}

### 2017-10-27 paper/sections/type-theory.tex
Cosmetic changes.

\begin{notation}
  Types are written using \abbre{small caps}.  We write the type
  judgments as $m : \abbre{m}$ to denote that~$m$ is of
  type~$\abbre{m}$.
\end{notation}

We produce derivation trees inside the judgments using the derivation
rules.  Since type theory can be seen as typed $λ$-calculus with
dependent products, evaluation of $λ$-terms also called
\emph{normalization} of the proof-terms, are reductions using the
system inference rules.

Therefore, the task to verify a proof that proves a certain theorem is
translated to verify if the associated tree of the proof has the
correspondent type of the theorem. We know this process as
\emph{type-checking}. This feature of type theory allow us to verify a
proof generated by an \ATP by reconstructing its proof-term to
type-check the proof. For such a purpose, we use a proof assistant in
type theory like \Agda to delegate this task.  However, this approach
has some weak points discussed in~\cite{Bezem2002}. Some limitations
from the type theory point of view for proof-reconstruction is
described in the following sections.

>>>>

\begin{notation}
  Types are written using \abbre{small caps}.  We write the type
  judgments as $a : \abbre{a}$ to denote that~$a$ is of
  type~$\abbre{a}$.
\end{notation}

We produce derivation trees inside the judgments using the derivation
rules.  Since type theory can be seen as typed $λ$-calculus with
dependent function types, evaluation of $λ$-terms also called
\emph{normalization} of the proof-terms, are reductions using the
system inference rules. Therefore, the task to verify a proof that
proves a certain theorem is translated to verify if the associated
tree of the proof has the correspondent type of the theorem. We know
this process as \emph{type-checking}. This feature of type theory
allow us to verify a proof generated by an \ATP by reconstructing its
proof-term to type-check the proof. For such a purpose, we use a proof
assistant in type theory like \Agda to delegate this task.  However,
this approach has some weak points discussed in~\cite{Bezem2002}. Some
limitations from the type theory point of view for
proof-reconstruction is described in the following sections.
